quicR: An R Library for Streamlined Data Handling of Real-Time Quaking Induced Conversion Assays

### 2025-02-12 man/manuscript/softwareX_submission.tex
Updated softwareX submission files.

\title{Title (quicR: An R Library for Streamlined Data Handling of Real-Time Quaking Induced Conversion Assays)}

\author[label1,label2,label3]{Gage Rowden}
\author[label1,label2,label3]{Peter Larsen}
\address[label1]{Department of Veterinary and Biomedical Sciences, University of Minnesota, St. Paul, MN 55108, USA.}
\address[label2]{Minnesota Center for Prion Research and Outreach, University of Minnesota, St. Paul, MN 55108, USA.}
\address[label3]{Priogen Corp., St. Paul, MN 55114, USA.}

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
R package \sep R library \sep RT-QuIC \sep prion \sep diagnostics \sep CJD \sep Parkinson's

@@

\begin{table}[ht]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Metadata} \\
\hline
C1 & Current code version & v2.1.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & For example: \url{https://github.com/gage1145/quicR} \\
\hline
C3  & Permanent link to Reproducible Capsule & For example: \url{https://codeocean.com/capsule/0270963/tree/v1}\\
\hline
C4 & Legal Code License   & All software and code must be released under one of the pre-approved licenses listed in the \href{https://www.elsevier.com/journals/softwarex/2352-7110/guide-for-authors}{Guide for Authors}, such as Apache License, GNU General Public License (GPL) or MIT License. Write the name of the license you’ve chosen here. \\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & R \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & \\
\hline
C8 & If available Link to developer documentation/manual & For example: \url{https://github.com/gage1145/quicR/blob/main/man/manuscript/manuscript.pdf} \\
\hline
C9 & Support email for questions & rowde002@umn.edu\\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{codeMetadata} 
\end{table}

% \textit{Optionally, you can provide information about the current executable
% software version filling in the left column of
% Table~\ref{executabelMetadata}. Please leave the first column as it is.}

\begin{table}[ht]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{(Executable) software metadata description} & \textbf{Please fill in this column} \\
\hline
S1 & Current software version & For example 1.1, 2.4 etc. \\
\hline
S2 & Permanent link to executables of this version  & For example: \url{https://github.com/combogenomics/DuctApe/releases/tag/DuctApe-0.16.4} \\
\hline
S3  & Permanent link to Reproducible Capsule & \\
\hline
S4 & Legal Software License & List one of the approved licenses \\
\hline
S5 & Computing platforms/Operating Systems & For example Android, BSD, iOS, Linux, OS X, Microsoft Windows, Unix-like , IBM z/OS, distributed/web based etc. \\
\hline
S6 & Installation requirements \& dependencies & \\
\hline
S7 & If available, link to user manual - if formally published include a reference to the publication in the reference list & For example: \url{http://mozart.github.io/documentation/} \\
\hline
S8 & Support email for questions & \\
\hline
\end{tabular}
\caption{Software metadata (optional)}
\label{executabelMetadata} 
\end{table}

\section{Motivation and significance}
\textit{In this section, we want you to introduce the scientific background and the motivation for developing the software.}

\section{Software description}

    Real-time quaking induced conversion (RT-QuIC) is a cutting-edge diagnostic assay that has garnered significant attention for its ultra-sensitive detection of misfolded protein aggregates \cite{Wilham2010, Atarashi2011}. The assay works by converting a recombinant protein substrate into an amyloid aggregate in the presence of a misfolded seed \cite{Wilham2010, Orru2012, Orru2017, Orru2015, Bongianni2019, Dassanayake2016, Hwang2018, Groveman2018, Metrick2020}. The assay's sensitivity and specificity make RT-QuIC a promising tool for diagnosing diseases such as prion disorders and other protein misfolding pathologies \cite{Fiorini2020, Franceschini2017, Picasso-Risso2022, Holz2021}. However, the relatively recent development and novelty of the assay have left a gap in widely accepted academic and industry standards for data analysis and interpretation \cite{Rowden2023}.

@@

\subsection{Software architecture}
    quicR was developed to address the growing need for efficient data conversion, analysis, and visualization of RT-QuIC data. With a focus on usability and reproducibility, the package is designed to standardize workflows and ensure compatibility across multiple laboratories. Its primary input format is data exported as Excel workbooks from the proprietary MARS software (BMG Labtech, Ortenberg, Germany), providing seamless integration with existing experimental workflows. See Figure \ref{fig:workflow} for a detailed workflow.
    \begin{figure}[ht]
        \caption{Workflow hierarchy of the quicR package. Blue nodes indicate steps where BMG software is needed. Purple nodes indicate functions dedicated to handling metadata. Red nodes are functions that acquire and manipulate raw data. Orange nodes are functions which calculate some metric. Finally, yellow nodes represent data analysis endpoints.}
        \centering
        \includegraphics[width=\textwidth]{images/workflow2.png}
        \label{fig:workflow}
    \end{figure}

>>>>

\title{quicR: An R Library for Streamlined Data Handling of Real-Time Quaking Induced Conversion Assays}

\author[label1,label2,label3]{Gage Rowden\corref{cor}\fnref{email}}
\ead{rowde002@umn.edu}
\fntext[email]{}
\cortext[cor]{}
\author[label1,label2,label3]{Peter Larsen}
\address[label1]{Department of Veterinary and Biomedical Sciences, University of Minnesota, St. Paul, MN 55108, USA.}
\address[label2]{Minnesota Center for Prion Research and Outreach, University of Minnesota, St. Paul, MN 55108, USA.}
\address[label3]{Priogen Corp., St. Paul, MN 55114, USA.}

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
R package \sep RT-QuIC \sep prion \sep diagnostics \sep CJD \sep Parkinson's

@@

\begin{table}[ht]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Metadata} \\
\hline
C1 & Current code version & V2.1.0 \\
\hline
C2 & Permanent link to code/repository used for this code version & \url{https://github.com/gage1145/quicR} \\
\hline
C3  & Permanent link to Reproducible Capsule & \url{https://github.com/gage1145/quicR/releases/tag/v2.1.0}\\
\hline
C4 & Legal Code License   & GPL-3 \\
\hline
C5 & Code versioning system used & git \\
\hline
C6 & Software code languages, tools, and services used & R \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & R (>=4.1.0) \\
\hline
C8 & If available Link to developer documentation/manual & \url{https://cran.r-project.org/web/packages/quicR/quicR.pdf} \\
\hline
C9 & Support email for questions & rowde002@umn.edu\\
\hline
\end{tabular}
\caption{Code metadata}
\label{codeMetadata} 
\end{table}

\section{Motivation and significance}
    Real-time quaking induced conversion (RT-QuIC) is a cutting-edge diagnostic assay that has garnered significant attention for its ultra-sensitive detection of misfolded protein aggregates \cite{Wilham2010, Atarashi2011}. The assay works by converting a recombinant protein substrate into an amyloid aggregate in the presence of a misfolded seed \cite{Wilham2010, Orru2012, Orru2017, Orru2015, Bongianni2019, Dassanayake2016, Hwang2018, Groveman2018, Metrick2020}. The assay's sensitivity and specificity make RT-QuIC a promising tool for diagnosing diseases such as prion disorders and other protein misfolding pathologies \cite{Fiorini2020, Franceschini2017, Picasso-Risso2022, Holz2021}. However, the relatively recent development and novelty of the assay have left a gap in widely accepted academic and industry standards for data analysis and interpretation \cite{Rowden2023}.

@@

\textit{In this section, we want you to introduce the scientific background and the motivation for developing the software.}

\section{Software description}
    \subsection{Software architecture}
        quicR was developed to address the growing need for efficient data conversion, analysis, and visualization of RT-QuIC data. With a focus on usability and reproducibility, the package is designed to standardize workflows and ensure compatibility across multiple laboratories. Its primary input format is data exported as Excel workbooks from the proprietary MARS software (BMG Labtech, Ortenberg, Germany), providing seamless integration with existing experimental workflows. See Figure \ref{fig:workflow} for a detailed workflow.
        \begin{figure}[ht]
            \caption{Workflow hierarchy of the quicR package. Blue nodes indicate steps where BMG software is needed. Purple nodes indicate functions dedicated to handling metadata. Red nodes are functions that acquire and manipulate raw data. Orange nodes are functions which calculate some metric. Finally, yellow nodes represent data analysis endpoints.}
            \centering
            \includegraphics[width=\textwidth]{images/workflow2.png}
            \label{fig:workflow}
        \end{figure}
